\thetitle

### 2021-04-30 writing/rational-natural.tex
Typo fixes
Extend definition of blame modes to use stack in absence of blame
And add a paragraph explaining why.

Here is a rigorous definition. 
\begin{quote}
\it A \emph{Natural blame trail} is a sequence of scenarios $\conf_0,...\conf_n$ of
a program $\system$ such that for all $0 \leq i \leq n - 1$, $\conf_i \subset
\conf_{i+1}$ and $\conf_{i+1} \setminus \conf_i =\{\blame{\system, \conf_i}\}$, where
$\blame{\system, \conf}$ denotes the module that $\conf$ blames.
\end{quote}

@@

To account for this situation, a new mode of the Natural rational
programmer follows a migration process based on exceptions.
\begin{quote}
\it A {\em Natural exception trail\/} is a sequence of scenarios $\conf_0,...\conf_n$ of a
program $\system$ such that for all $0 \leq i \leq n - 1$, $\conf_i \subset
\conf_{i+1}$ and $\conf_{i+1} \setminus \conf_i = \{\exception{\system, \conf_i}\}$
where $\exception{\system, \conf}$ denotes the location of the exception of $\conf$.
\end{quote}

>>>>

Here is a rigorous definition. 
\begin{quote}
\it A \emph{Natural blame trail} is a sequence of scenarios $\conf_0,...\conf_n$ of
a program $\system$ such that for all $0 \leq i \leq n - 1$, $\conf_i \subset
\conf_{i+1}$ and
\[ \conf_{i+1} \setminus \conf_i = %
\left\{ %
\begin{tabular}{ll} %
$\{\blame{\system, \conf_i}\}$      & if $\conf_i$ produces blame \\ %
$\{\exception{\system, \conf_i}\}$  & otherwise \\ %
\end{tabular} %
\]
where $\blame{\system, \conf}$ denotes the module (of $\system$) that $\conf$ blames,
and $\exception{\system, \conf}$ denotes the first untyped location in the stacktrace produced by $\conf$.
\end{quote}

This definition rests on the two options for $\conf_{i+1} \setminus \conf_i$.
The first part reflects the rational programmer's use of blame to extend the
trail. However, blame may not be available in some scenarios. The second part
accounts for those scenarios that produce errors from the runtime system before
any impedance mismatch can be detected. For example, the fully untyped
configuration can only produce such an error (e.g. from {\tt +} receiving a
string). Exceptions from the runtime do not carry blame, so
the rational programmer proceeds using the accompanying stacktrace instead.

@@

To account for this situation, a new mode of the Natural rational
programmer follows a migration process based on entirely on exceptions.
\begin{quote}
\it A {\em Natural exception trail\/} is a sequence of scenarios $\conf_0,...\conf_n$ of a
program $\system$ such that for all $0 \leq i \leq n - 1$, $\conf_i \subset
\conf_{i+1}$ and $\conf_{i+1} \setminus \conf_i = \{\exception{\system, \conf_i}\}$.
\end{quote}

### 2021-04-30 writing/rational-transient.tex
Extend definition of blame modes to use stack in absence of blame
And add a paragraph explaining why.

\begin{quote}
\it A \emph{Transient-first blame trail}/\emph{Transient-last blame trail} is a sequence of scenarios
$\conf_0,...\conf_n$ of $\system$ where for all $0 \leq i \leq n - 1$,
$\conf_i \subset \conf_{i+1}$ and $\conf_{i+1} \setminus \conf_i =
\first{\mblame{\system, \conf_i}}$ where $\first{\mblame{\system,
\conf}}$/$\last{\mblame{\system, \conf}}$ is the
first/last module, respectively,  that Transient adds to the blame sequence for $\conf$.
\end{quote}

% \begin{quote}
% \it A  is a sequence of scenarios
% $\conf_0,...\conf_n$ of $\system$ where for all $0 \leq i \leq n - 1$,
% $\conf_i \subset \conf_{i+1}$ and $\conf_{i+1} \setminus \conf_i =
%   \last{\mblame{\system, \conf_i}}$ where $\last{\mblame{\system, \conf}}$ is the
% last module Transient adds to the blame sequence for $\conf$.
% \end{quote}

>>>>

\begin{quote}
\it A \emph{Transient-first blame trail}/\emph{Transient-last blame trail} is a sequence of scenarios
$\conf_0,...\conf_n$ of $\system$ where for all $0 \leq i \leq n - 1$,
$\conf_i \subset \conf_{i+1}$ and
\[ \conf_{i+1} \setminus \conf_i = %
\left\{ %
\begin{tabular}{ll} %
$\{\first{\mblame{\system, \conf_i}}\}$  & if $\conf_i$ produces blame \\ %
$\{\exception{\system, \conf_i}\}$     & otherwise \\ %
\end{tabular} %
\]
where $\first{\mblame{\system, \conf}}$/$\last{\mblame{\system, \conf}}$ is the
first/last module, respectively,  that Transient adds to the blame sequence for $\conf$.
\end{quote}
